Mendelian randomization for causal inference in genome-wide association studies

### 2022-03-01 mr-review.Rmd
mrcip text added

Likelihood-based MR methods, unlike two-stage methods, provide maximum likelihood 
estimates with their many desirable properties [@fisher1925theory]. Limited information maximum likelihood from econometrics 
is the earliest approach for likelihood-based inference in MR [@davidson1993estimation].

@@

>>>>

Likelihood-based MR methods, unlike two-stage methods, provide maximum likelihood 
estimates with their many desirable properties [@fisher1925theory]. Limited information maximum likelihood from econometrics 
is the earliest approach for likelihood-based inference in MR [@davidson1993estimation;@hansen1996finite].

@@

@jiang_constrained_2019, working with individual-level GWAS data, share naive and smoothed constrained instrumental variable methods. Central to their
work is the proposal to create a new instrumental variable as a linear combination of genotype data from 
a collection of SNP instruments. They require that the new instrumental variable be standardized to have norm 1 and be orthogonal to potentially vertically  
pleiotropic traits. For a collection of $K$ SNPs, they choose the $K$-vector that meets these criteria and 
maximizes the correlation with the exposure. To obtain their second estimator, @jiang_constrained_2019 apply a 
$\ell_0$ penalty to constrain the number of SNP instruments that are given nonzero weights in the calculation 
of the new instrumental variable. They then compare their method with sisVIVE and allele score methods. 
@jiang_constrained_2019 find that both the smoothed version of their estimator and the allele score method are 
unbiased in their simulation settings.

@xu_mrcip_2021 consider two classes of horizontal pleiotropy with their method, MRCIP. They partition horizontal 
pleiotropic effects into correlated or idiosyncratic horizontal pleiotropy. Idiosyncratic horizontal pleiotropy 
is that with a large effect size, while correlated horizontal pleiotropy, like its definition elsewhere in our 
manuscript, refers to correlation between the effects of the SNP instruments on the exposure and the effects of 
the SNP instruments on the outcome. MRCIP relaxes both the exclusion restriction assumption and the InSIDE 
assumption by directly modeling the correlated horizontal pleiotropy with random effects. Furthermore, to 
accommodate the idiosyncratic pleiotropy, the authors downweight SNP instruments that demonstrate strong direct 
effects on the outcome. The find that their MRCIP provides valid causal inference even when there are no valid 
SNP instruments (*i.e.*, when all SNP instruments are invalid) and when the InSIDE assumption is violated 
[@xu_mrcip_2021]. Their PRW-EM algorithm adds a reweighting step to the traditional expectation maximization algorithm to estimate the causal effect parameter in MR analysis [@dempster1977maximum]. Finally, they compare  MRCIP to other MR methods with both simulated and real data sets. @xu_mrcip_2021, like the developers of MR-RAPS, assume that the horizontal pleiotropy is balanced, *i.e.*, that the mean horizontal pleiotropic effect is zero. One direction for future study is to augment MRCIP to permit unbalanced horizontal pleiotropy.
